Automated Example Oriented REST API Documentation

### 2017-09-24 introduction.tex
added changes based on feedback

Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}. Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}. APIs enable multiple software components to interact with each other. Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism of information between systems.

REST APIs are a type of API that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. He described the design of modern Web as an example of a REST-style architecture. REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by levergading a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.

REST APIs allow access to application resources over HTTP. While REST describes a generic protocol for accessing application resources, the applications are responsible for the specific representation of the resources. For example, a REST API may support multiple Internet media types such as JSON, XML, CSV for a single application resource. Similarly, two diffenent REST APIs may use the same media type but contain entirely different formats to represent their objects.

@@

\section{Research Goals}
The overarching goal of my research is to improve the current state of the proces of REST API documentation process by automating the repeating manual steps in the aforementioned workflow. More specifically, I aimed to solve the following research goals:

@@

\begin{itemize}
  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background for improving the state of REST API documentation process.
  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitiative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
\end{itemize}

@@

\begin{enumerate}
  \item A list of REST API documentation requirements derived from the existing literature and the current state of industry practices to be used as a guideline by practitioners and researchers.
  \item SpyREST, a novel technique and a reusable tool that uses interception to automatically generate and maintain evolving REST API documentation with usage examples from executable code.
  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that practitioners can follow to improve their REST API documentation experience.
  \item An empirical evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
\end{enumerate}

>>>>

Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}. Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}. APIs enable multiple software components to interact with each other. Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism between systems.

REST APIs are a type of API that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. He described the design of modern Web as an example of a REST-style architecture. REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by leveraging a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.

REST APIs allow access to application resources over HTTP. While REST describes a generic protocol for accessing application resources, the applications are responsible for the specific representation of the resources. For example, a REST API may support multiple Internet media types such as JSON, XML, CSV for a single application resource. Similarly, two different REST APIs may use the same media type but contain entirely different formats to represent their objects.

@@

\section{Research Goals}
The overarching goal of my research is to improve the current state of the process of REST API documentation process by automating the repeating manual steps in the aforementioned workflow. More specifically, I aimed to solve the following research goals:

@@

\begin{itemize}
  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background for improving the state of REST API documentation process.
  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
\end{itemize}

@@

\begin{enumerate}
  \item A list of REST API documentation requirements derived from the existing literature and the current state of industry practices to be used as a guideline by practitioners and researchers.
  \item SpyREST, a novel technique and a reusable tool that uses interception to automatically generate and maintain evolving REST API documentation with usage examples from executable code.
  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that REST API developers can follow to improve their REST API documentation experience.
  \item An empirical evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
\end{enumerate}

### 2017-09-24 root.tex
added changes based on feedback

%
%                 Substitute  List of Examiners
%
\begin{signing}{Department of Computer Science}
\signline
Dr. Frank Maurer \\
Supervisor\\
Department of Computer Science \\
University of Calgary \\
Canada\\
\signline
Dr. Mea Wang\\
Supervisory Committee Member\\
Department of Computer Science \\
University of Calgary \\
Canada\\
\signline
Dr. Craig Anslow\\
Supervisory Committee Member\\
School of Eng. and Comp. Science \\
Victoria University of Wellington\\
New Zealand\\
\end{signing}
%
\newpage
\phantomsection
\altchapter{\bf{Abstract}}
API documentation presents both a problem and an opportunity for API usability. REST APIs provide interconnectivity between applications over HTTP. Documentation of a REST API is a key information source for API client developers. Most REST APIs are documented using a manual approach, which can be time consuming and error-prone. REST API developers need to efficiently document their APIs with qualities that make the API usable. In this research, I focused on the topic of automated REST API documentation to satisfy this need.

In this thesis, I present and evaluate of a novel technique to solve REST API documentation requirements. I present a set of REST API documentation requirements by studying the existing literature and the current industry practice. From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation. I present, SpyREST, a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples. The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation. I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use. From this study, I found that it was feasible to leverage API test code to automatically generate an always-updated REST API documentation for evolving REST APIs with usage examples. To evaluate the impact of usage examples on REST API client developers, I performed a controlled study with experienced REST API developers. From this case study, I found that REST API client developers faced patterns of obstacles using REST APIs that can be reduced by including API usage examples in the API documentation, as suggested in the proposed technique.

>>>>

%
%                 Substitute  List of Examiners
%
\begin{signing}{Department of Computer Science}
\signline
Dr. Frank Maurer \\
Supervisor\\
Department of Computer Science \\
University of Calgary, Canada \\
\signline
Dr. Mea Wang\\
Supervisory Committee Member\\
Department of Computer Science \\
University of Calgary, Canada \\
\signline
Dr. Craig Anslow\\
Supervisory Committee Member\\
School of Engineering and Computer Science \\
Victoria University of Wellington, New Zealand\\
\signline
Dr. Nils Daniel Forkert\\
Internal Examiner\\
Department of Radiology and Hotchkiss Brain Institute\\
University of Calgary, Canada\\
\signline
Dr. Yves Lucet\\
External Examiner\\
Computer Science\\
University of British Columbia (Okanagan), Canada\\
\end{signing}
%
\newpage
\phantomsection
\altchapter{\bf{Abstract}}
API documentation presents both a problem and an opportunity for API usability. REST APIs provide interconnectivity between applications over HTTP. Documentation of a REST API is a key information source for API client developers. Most REST APIs are documented using a manual approach, which can be time consuming and error-prone. REST API developers need to efficiently document their APIs with qualities that make the API usable. In this research, I focused on the topic of automated REST API documentation to satisfy this need.

In this thesis, I present and evaluate of a novel technique to solve REST API documentation requirements. I present a set of REST API documentation requirements by studying the existing literature and the current industry practice. From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation. I present, \textit{SpyREST}, a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples. The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation. I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use. From this study, I found that it was feasible to leverage API test code to automatically generate an always-updated REST API documentation for evolving REST APIs with usage examples. To evaluate the impact of usage examples on REST API client developers, I performed a controlled study with experienced REST API developers. From this case study, I found that REST API client developers faced patterns of obstacles using REST APIs that can be reduced by including API usage examples in the API documentation, as suggested in the proposed technique.
